---
title: "KalmanFilterStudy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Anlysis for Nile
Trial

## Pre-analysis

### Time series plot
```{r plot}
plot(Nile, main = "Time series plot")
```


### Histgram
```{r hist}
hist(Nile, main="histgram", xlab="values")
```


### Summary
```{r summary}
summary(Nile)
```


### autocorrelations
```{r acf}
par(mar=c(4,4,3,1))
acf(Nile, main="auto correltion")
```


### Spectual analysis
``` {r spectal}
source("plot_spectrum.R")
plot.spectrum(Nile, lab="Year", main="spectrum")
```


### Holt-Winters
``` {r Holt-Winters}
HW_Nile <- HoltWinters(Nile, beta=FALSE, gamma=FALSE)
mygray <- "#80808080"
plot(HW_Nile, main="Holt-Winters",
     col=mygray, col.predicted="black", lty.predicted="dashed")
```


### Holt-Winters decomposed
``` {r Holt-Winters decomp}
HW_out <- HW_Nile
HW_decomp <- ts.union(
  y = HW_out$x,
  Level = HW_out$fitted[, "level"],
  Residuals = residuals(HW_out)
)
plot(HW_decomp, main = "Holt-Winters decomp")
```


### Autocorrelation for residuals
```{r autocorrel for residuals}
par(mar=c(4,4,3,1))
acf(residuals(HW_Nile), main = "autocorrel for residuals")
```


## Modelling by Kalman filter
used local level model.


### Preprocess: specify model parameters
```{r preprocess, echo = FALSE}
library(dlm)
```

```{r modelling}
mod <- dlmModPoly(order = 1)
build_dlm <-function(par) {
  mod$W[1, 1] <- exp(par[1])
  mod$V[1, 1] <- exp(par[2])
  return(mod)
}

lapply(list(c(0, 0), c(1, 10), c(20, 3)),
       function(params) {
         dlmMLE(y = Nile, parm = params, build = build_dlm)
       })
```

```{r param estimation by MLE}
fit_dlm <- dlmMLE(y = Nile, parm = c(0, 0),
                  build =  build_dlm,
                  hessian = TRUE)
```

* calculate (asymptotic) std error from Hessian by delta method
```{r mle}
exp(fit_dlm$par) * sqrt(diag(solve(fit_dlm$hessian)))
```

```{r set model}
mod <- build_dlm(fit_dlm$par)
mod
```


### Kalman filtering

```{r filtering}
dlm_filtering_obj <- dlmFilter(y = Nile, mod = mod)
str(dlm_filtering_obj, max.level = 1)
```

```{r calc mean and stddev of filtering dist}
m <- dropFirst(dlm_filtering_obj$m)
m_sdev <- sqrt(
  dropFirst(as.numeric(
    dlmSvd2var(dlm_filtering_obj$U.C, dlm_filtering_obj$D.C)
  ))
)
```

```{r calc quantile of filtering dist}
m_quant <- list(m + qnorm(0.025, sd = m_sdev),
                m + qnorm(0.975, sd = m_sdev))
```

```{r plot result of filtering}
ts.plot(cbind(Nile, m, do.call("cbind", m_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))
legend(legend = c("observed", "mean of filtered", "95% interval of filtered"),
       col = c("lightgray", "black", "black"),
       lty = c("solid", "solid", "dashed"),
       x = "topright", text.width = 24, cex= 0.6
       )
```

### kalman forecasting
```{r forecast}
dlm_forecast_obj <- dlmForecast(
  mod = dlm_filtering_obj,
  nAhead = 10,
  sampleNew = 10
)
str(dlm_forecast_obj, max.level = 1)
```


### calc mean and stddev of predict dist
```{r calc mean and stddev of predict dist}
a <- ts(data = dlm_forecast_obj$a, start = c(1971, 1))
a_sdev <- sqrt(as.numeric(
  dlm_forecast_obj$R
))
```


### calc quantile of predict dist
```{r calc quantile of predict dist}
a_quant <- list(a + qnorm(0.025, sd = a_sdev),
                a + qnorm(0.975, sd = a_sdev))
```


### plot result of forecast
```{r plot forecast}
ts.plot(cbind(Nile, a, do.call("cbind", a_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))
legend(legend = c("observed", "mean of forecast", "95% interval of forecast"),
       col = c("lightgray", "black", "black"),
       lty = c("solid", "solid", "dashed"),
       x = "topright", text.width = 28, cex= 0.6
       )
```


### Kalman smoothing
```{r smoothing}
dlm_smoothing_obj <- dlmSmooth(
  y = Nile,
  mod = mod
)
str(dlm_smoothing_obj, max.level = 1)
```


### calc mean and stddev of smoothed dist
```{r calc mean and stddev of smoothed dist}
s <- dropFirst(dlm_smoothing_obj$s)
s_sdev <- sqrt(dropFirst(
  as.numeric(
    dlmSvd2var(dlm_smoothing_obj$U.S,
               dlm_smoothing_obj$D.S)
  ))
)
```


### calc quantile of smoothed dist
```{r calc quantile of smoothed dist}
s_quant <- list(s + qnorm(0.025, sd = s_sdev),
                s + qnorm(0.975, sd = s_sdev))
```


### plot result of smoothed
```{r plot smoothed}
ts.plot(cbind(Nile, s, do.call("cbind", s_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))
legend(legend = c("observed", "mean of forecast", "95% interval of forecast"),
       col = c("lightgray", "black", "black"),
       lty = c("solid", "solid", "dashed"),
       x = "topright", text.width = 28, cex= 0.6
       )
```


## Assesment of model

### Likelihood
```{r likelihood}
dlmLL(y = Nile, mod = mod)
```

### Innovation
```{r check autocorrelation for filtered}
par(mar = c(4, 4, 3, 1))
tsdiag(object = dlm_filtering_obj)
```

```{r check normalization}
e <- residuals(object = dlm_filtering_obj, sd = FALSE)
qqnorm(e)
qqline(e)
```
